---
title: "Tutorial"
author: "Edgar Treischl"
date: "01.01.2020"
output: 
  learnr::tutorial:
    highlight: kate
    theme: flatly
    
description: >
  Learn about the base data types in R. Explore R's data frames, and learn how
  to interact with data frames and their columns.
   
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(ggplot2)
library(tidyverse)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)

```

## The Basics

This tutorial summarizes the five main functions from the `dplyr` package which help you to handle most of the common data manipulation tasks.

-   *Learn how to pick* cases depending on their values with `filter()`
-   *Sort* the data with `arrange()`
-   *Choose* variables based on their names with `select()`
-   *Extend* the dataset and generate new variables with `mutate()`
-   *Summarize* variables with `summarise()`




## Filter()

Use `filter()` to subset the data based on specific values. Dplyr filters the `mtcars` data depending on the specified conditions and returns a new data frame. Adjust the filter and use boolean operations to combine several conditions or use relational operators to select observations.

### Practice

1.  Adjust the code so that only cars with an manual transmission (stored in the variable am: 0 = automatic, 1 = manual) are in the dataframe.

```{r filter1, exercise=TRUE, exercise.lines = 5, exercise.eval=FALSE}
filter(mtcars, hp > 100)
```

```{r filter1-hint}
#Did you mean `am == 0`?
```

2.  Create a dataframe that include all observations, but without cars that have 8 cylinders (cyl: Number of cylinders).

```{r filter2, exercise=TRUE, exercise.lines = 5, exercise.eval=FALSE}
filter(mtcars, hp > 100)
```

```{r filter2-hint}
#Did you mean `cyl == !8`?
```

3.  Can you adjust the code so that two condition have to be fullfilled simultaneously. For example, choose cars with more than 100 hp and less then 200 hp.

```{r filter3, exercise=TRUE, exercise.lines = 5, exercise.eval=FALSE}
filter(mtcars, hp > 100)
```

```{r filter3-hint}
#filter(mtcars, hp > 100 & hp < 200)
```


## Arrange()

Sometimes it is handy if data is sorted in a specific way. You can sort or arrange the order of your data frame with `arrange()`. For example, we can arrange the cars from the lowest to the highest value of horsepower, can you stille remember how to sort in a descending order?

```{r filter4, exercise=TRUE, exercise.lines = 2}
arrange(mtcars, hp)
```

```{r filter4-hint}
#Use `desc(variablename)` to order the data frame in descending order :-)
```

## Select ()

In this tutorial we use a rather small dataset. In real life, you will encounter messy and large datasets. Use `select()` to specify which variables (e.g. mpg, hp) we want to keep from the `mtcars` data. 


1.  Select all columns from the mtcars data between horsepower (hp) and automatic transmission (am).

```{r filter5, exercise=TRUE, exercise.lines = 5, exercise.eval=FALSE}
select(mtcars, ...) %>% slice(1:5)
```

```{r filter5-hint, exercise.lines = 2}
select(mtcars, (mpg:hp)) %>% slice(1:5)
```

2.  Select all variables which start with a "d".

```{r filter6, exercise=TRUE, exercise.lines = 5, exercise.eval=FALSE}
select(mtcars, ...) %>% slice(1:5)
```

```{r filter6-hint, exercise.lines = 2}
select(mtcars, starts_with("d")) %>% slice(1:5)
```

3.  Select all variables which end with a "p".

```{r filter7, exercise=TRUE, exercise.lines = 5, exercise.eval=FALSE}
select(mtcars, ...) %>% slice(1:5)
```

```{r filter7-hint, exercise.lines = 2}
select(mtcars, ends_with("p")) %>% slice(1:5)
```


## Mutate ()

Often, the raw data does not contain the variables in the way we need them. In such an instance we have to transform the data, generate new variables depending on the values of other variables in the dataset. Here comes `mutate()` into play, which adds a generated variable (or column) at the end of the data. 



### Practice

1.  The data set contains the weight (`wt`) of the cars saved in pounds (1000 lbs). Can you transfer the pound in kilograms (hint: `0.453592`)? Maybe the code from the Chapter helps you.

```{r filter7b, exercise=TRUE, exercise.lines = 7, exercise.eval=FALSE}
select(mtcars, hp, gear) #a subset of the data set
mutate(df,
       kw = hp * 0.74570,
       hp2 = kw * 1.34102,
       hp_rounding = round(hp2, 1)
) %>% slice(1:5)
```

```{r filter7b-hint, exercise.lines = 2}
select(mtcars, wt) |> 
  mutate(wt_new = wt * 0.453592) |>  
  slice(1:5)
```

2.  Check whether the transformation is right! Reverse the calcuation and save the results in the second variable (hint: `2.205`). Use the `round()` function, but don't forget to adjust it for the weight variable.

```{r filter8, exercise=TRUE, exercise.lines = 7, exercise.eval=FALSE}
#Enter code here ;)
```

```{r filter8-hint, exercise.lines = 2}
select(mtcars, wt) |> 
  mutate(wt_new = wt/0.453592,
         reverse = round(wt_new/2.205, 3)) |>  
  slice(1:5)
```

3.  Generate a logical variable that indicate whether the transmission of a car is automatic.

```{r filter9, exercise=TRUE, exercise.lines = 7, exercise.eval=FALSE}
###
```

```{r filter9-hint, exercise.lines = 6}
mtcars |> 
  select(am) |> 
  mutate(automatic = ifelse(am == 0, "TRUE", "FALSE"))|> 
  slice(1:5)
```



## Summarise()

We can calculate measurements of central tendencies (e.g. means, modus) with the `summarise()` function, which collapses several columns into a single row. 



1.  Calculate the mean horsepower `hp` for each group of transmission `am`.  

```{r filter10, exercise=TRUE, exercise.lines = 10, exercise.eval=FALSE}
#adjust me!
```

```{r filter10-solution, exercise.lines = 6}
mtcars %>% 
  group_by(am) %>% 
  summarise(
    mean_hp = mean(hp)
  )
```

2. Make sure that your summary does not rely on an unlawful small number of observations. To get the number of observation per group, add `count = n()` to the summary function.

```{r filter11, exercise=TRUE, exercise.lines = 10, exercise.eval=FALSE}
#...
```

```{r filter11-solution, exercise.lines = 6}
mtcars %>% 
  group_by(am) %>% 
  summarise(
    mean_hp = mean(hp),
    count = n()
  )
```

3.  Calculate the `median(x)` instead of the mean and add the standard deviation (`sd(x)`).

```{r filter12, exercise=TRUE, exercise.lines = 10, exercise.eval=FALSE}
#...
```

```{r filter12-solution, exercise.lines = 6}
mtcars %>% 
  group_by(am) %>% 
  summarise(
    median_hp = median(hp),
    sd_hp = sd(hp)
  )

```



## The End

Congratulations, mission accomplished! You are now able to *handle and manipulate* data, which gives you the skills for your own analysis. 







