---
title: "PracticeR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PracticeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `PracticeR` package gives access to tutorials, templates, and further materials of the Practice R book. First of all, it give access to source code of the book. The `show_script()` function returns the source code for each chapter. It copies the code, opens a new R script and inserts the code for the following chapters.

```{r, error=TRUE}
library(PracticeR)
show_script()
```

In addition, the package includes some convenient functions to learn R. For example, the `show_linetypes` function returns the line types graph.

```{r}
show_linetypes()
```

The `show_shapetypes` function returns the shape types graph.

```{r}
show_shapetypes()
```

And the `transformer()` function transforms a numerical variable according to the ladder of powers and returns histograms to inspect them.

```{r}
transformer(mtcars$mpg)
```

